Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-07-23 exp1/iswc2013.tex
smaller font in code snippets
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@608 43ef1190-90fd-4dd7-a769-d0121f02b80e
figure placement and sizing
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@609 43ef1190-90fd-4dd7-a769-d0121f02b80e

\documentclass{llncs}
%\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{cite}
\usepackage{verbatim}
\usepackage{graphicx}

@@

\begin{tabular}{p{0.45\textwidth}p{0.45\textwidth}}
Level 1 & Level 2 \\ 
\verbatiminput{experiment/BGPComp-1} & \verbatiminput{experiment/BGPComp-2} \\
\end{tabular}

@@

\begin{tabular}{p{0.40\textwidth}p{0.55\textwidth}}

``Union'' & ``Lang'' \\
\verbatiminput{experiment/Union-2} &
\verbatiminput{experiment/Lang-2} \\

``Range'' & ``Optional'' \\
\verbatiminput{experiment/Range-2} & 
\verbatiminput{experiment/Optional-2}
\end{tabular}

These fragments has been carefully designed for illustrative utility,
as well as suitable selectivity. They are not themselves important,
they only serve to illustrate certain examples of possible
factors. When the experiment is run, they are combined to yield a
query, which is then sent to a SPARQL endpoint. The SPARQL endpoint is
itself set up using 4store (see \cite{harris20094store}) version
1.1.5. The ``Machine'' factor specifies a hostname, whereas
``TripleC'' and ``Implement'' specifies a port number. On each host,
four 4store instances runs sequentially and independently on different
ports, where processes that are not actively answering a query are
idle.

@@

\begin{figure}[h!]
  \includegraphics[width=\textwidth]{fullnormal.pdf}
  \caption{A normal plot of the Full Factorial Experiment. The
    labelled points are considered significant by using the Lenth
    criterion at a level $\alpha=0.05$.}\label{fig:fullnormal}
\end{figure}

@@

% latex table generated in R 2.15.1 by xtable 1.5-6 package
% Fri May 10 16:31:49 2013
% Modified by hand to remove 1's
\begin{table}[hb]
\begin{center}
\caption{The magnitude of effects for some important main effects and interactions.}\label{tab:effects}
\begin{tabular}{rr}
  \hline
Factors & Effect  \\ 
  \hline
Implement & -21.27 \\ 
  Implement:Optional & -11.49 \\ 
  Implement:Union & -6.21 \\ 
  Implement:TripleC:Lang1 & 4.12 \\ 
  TripleC:Lang & 4.20 \\ 
  Implement:TripleC & 4.32 \\ 
  Implement:BGPComp & 4.89 \\ 
  Lang:Union & 5.16 \\ 
  Implement:BGPComp:Lang & 5.16 \\ 
  BGPComp:Lang & 5.25 \\ 
  Implement:Lang:Union & 5.75 \\ 
  TripleC & 7.06 \\ 
  Implement:Lang & 7.73 \\ 
  Lang & 8.62 \\ 
  Optional & 11.30 \\ 
  Union & 17.18 \\ 
   \hline
\end{tabular}
\end{center}
\end{table}

@@

\begin{table}[ht]
\begin{center}
\caption{$p$-values for different parts of the experiment}\label{tab:pvaluesfull}
\begin{tabular}{cccccl}
  \hline
``TripleC'' & ``BGPComp'' & ``Lang'' & ``Union'' & ``Optional'' & $p$ \\ 
  \hline
  1 & 1 & 1 & 1 & 1 & 0.012 \\ 
  1 & 1 & 1 & 1 & 2 & $1.4 \cdot 10^{-09}$ \\ 
  1 & 1 & 1 & 2 & 1 & $3.1 \cdot 10^{-09}$ \\ 
  1 & 1 & 1 & 2 & 2 & $6.1 \cdot 10^{-11}$ \\ 
  1 & 1 & 2 & 1 & 1 & $3.3 \cdot 10^{-06}$ \\ 
  1 & 1 & 2 & 1 & 2 & $2.7 \cdot 10^{-09}$ \\ 
  1 & 1 & 2 & 2 & 1 & $2 \cdot 10^{-06}$ \\ 
  1 & 1 & 2 & 2 & 2 & $3.6 \cdot 10^{-10}$ \\ 
  1 & 2 & 1 & 1 & 1 & 0.014 \\ 
  1 & 2 & 1 & 1 & 2 & $1.2 \cdot 10^{-10}$ \\ 
  1 & 2 & 1 & 2 & 1 & $2.8 \cdot 10^{-14}$ \\ 
  1 & 2 & 1 & 2 & 2 & $4.1 \cdot 10^{-15}$ \\ 
  1 & 2 & 2 & 1 & 1 & $2.1 \cdot 10^{-05}$ \\ 
  1 & 2 & 2 & 1 & 2 & $2.7 \cdot 10^{-07}$ \\ 
  1 & 2 & 2 & 2 & 1 & 0.0072 \\ 
  1 & 2 & 2 & 2 & 2 & $1.6 \cdot 10^{-05}$ \\ 
  2 & 1 & 1 & 1 & 1 & 0.28 \\ 
  2 & 1 & 1 & 1 & 2 & $3 \cdot 10^{-07}$ \\ 
  2 & 1 & 1 & 2 & 1 & $3.3 \cdot 10^{-07}$ \\ 
  2 & 1 & 1 & 2 & 2 & $1.7 \cdot 10^{-08}$ \\ 
  2 & 1 & 2 & 1 & 1 & 0.0023 \\ 
  2 & 1 & 2 & 1 & 2 & $6.5 \cdot 10^{-07}$ \\ 
  2 & 1 & 2 & 2 & 1 & 0.00032 \\ 
  2 & 1 & 2 & 2 & 2 & $1.3 \cdot 10^{-06}$ \\ 
  2 & 2 & 1 & 1 & 1 & 0.013 \\ 
  2 & 2 & 1 & 1 & 2 & $1 \cdot 10^{-11}$ \\ 
  2 & 2 & 1 & 2 & 1 & $3.8 \cdot 10^{-11}$ \\ 
  2 & 2 & 1 & 2 & 2 & $4.1 \cdot 10^{-15}$ \\ 
  2 & 2 & 2 & 1 & 1 &   1 \\ 
  2 & 2 & 2 & 1 & 2 & $2.3 \cdot 10^{-05}$ \\ 
  2 & 2 & 2 & 2 & 1 &   1 \\ 
  2 & 2 & 2 & 2 & 2 & 0.99 \\ 
   \hline
\end{tabular}
\end{center}
\end{table}

@@

\begin{figure}[h!]
  \includegraphics[width=\textwidth]{frac32normal.pdf}
  \caption{A normal plot of the Fractional Factorial Experiment with
    32 runs. The labelled points are considered significant by using
    the Lenth criterion at a level
    $\alpha=0.05$.}\label{fig:frac32normal}
\end{figure}

@@

\begin{figure}[ht]
  \includegraphics[width=\textwidth]{frac64normal.pdf}
  \caption{A normal plot of the Fractional Factorial Experiment with
    64 runs. The labelled points are considered significant by using
    the Lenth criterion. Triangles correspond to a level $\alpha=0.15$
    and crosses at $\alpha=0.05$. The use of different symbols have
    been added to FrF2 by the authors.}\label{fig:frac64normal}
\end{figure}

@@

\begin{figure}[h!]
  \includegraphics[width=\textwidth]{frac64hugenormal.pdf}
  \caption{A normal plot of the Fractional Factorial Experiment with
    64 runs where the level 2 of ``TripleC'' is 15 MTriples. The
    labelled points are considered significant by using the Lenth
    criterion at a level $\alpha=0.05$.}\label{fig:frac64hugenormal}
\end{figure}

>>>>

\documentclass{llncs}
%\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{cite}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{wrapfig}

@@

  \begin{center}\small
    \begin{tabular}{|@{~~}p{0.40\textwidth}|@{~~}p{0.53\textwidth}|}
      \hline
      \textbf{Level 1} & \textbf{Level 2} \\[-2ex] 
      \verbatiminput{experiment/BGPComp-1}
      & \verbatiminput{experiment/BGPComp-2}\\
      \hline
    \end{tabular}
  \end{center}

@@

\begin{center}\small
  \begin{tabular}{|@{~~}p{0.40\textwidth}|@{~~}p{0.53\textwidth}|}
    \hline
    \textbf{``Union''} & \textbf{``Lang''} \\[-2ex]
    \verbatiminput{experiment/Union-2} &
    \verbatiminput{experiment/Lang-2} \\
    \hline
    \textbf{``Range''} & \textbf{``Optional''} \\[-2ex]
    \verbatiminput{experiment/Range-2} & 
    \verbatiminput{experiment/Optional-2}\\
    \hline
  \end{tabular}
\end{center}
These fragments has been carefully designed for illustrative utility,
as well as suitable selectivity. They are not themselves important,
they only serve to illustrate certain examples of possible
factors. When the experiment is run, they are combined to yield a
query, which is then sent to a SPARQL endpoint. The SPARQL endpoint is
itself set up using 4store (see \cite{harris20094store}) version
1.1.5. The ``Machine'' factor specifies a hostname, whereas
``TripleC'' and ``Implement'' specifies a port number. On each host,
four 4store instances runs sequentially and independently on different
ports, where processes that are not actively answering a query are
idle.

@@

\begin{figure}[t]
  \centerline{%
  \includegraphics[width=.9\textwidth]{fullnormal.pdf}}
  \caption{A normal plot of the Full Factorial Experiment. The
    labelled points are considered significant by using the Lenth
    criterion at a level $\alpha=0.05$.}\label{fig:fullnormal}
\end{figure}

@@

% latex table generated in R 2.15.1 by xtable 1.5-6 package
% Fri May 10 16:31:49 2013
% Modified by hand to remove 1's
\begin{wraptable}{l}{0.5\textwidth}
\caption{The magnitude of effects for some important main effects and interactions.}\label{tab:effects}
\medskip
\begin{tabular}{rr}
  \hline
Factors & Effect  \\ 
  \hline
Implement & $-$21.27 \\ 
  Implement:Optional & $-$11.49 \\ 
  Implement:Union & $-$6.21 \\ 
  Implement:TripleC:Lang1 & 4.12 \\ 
  TripleC:Lang & 4.20 \\ 
  Implement:TripleC & 4.32 \\ 
  Implement:BGPComp & 4.89 \\ 
  Lang:Union & 5.16 \\ 
  Implement:BGPComp:Lang & 5.16 \\ 
  BGPComp:Lang & 5.25 \\ 
  Implement:Lang:Union & 5.75 \\ 
  TripleC & 7.06 \\ 
  Implement:Lang & 7.73 \\ 
  Lang & 8.62 \\ 
  Optional & 11.30 \\ 
  Union & 17.18 \\ 
   \hline
\end{tabular}
\end{wraptable}

@@

\begin{table}[h!]
\begin{center}
\caption{$p$-values for different parts of the experiment}\label{tab:pvaluesfull}
\begin{tabular}{cccccl}
  \hline
``TripleC'' & ``BGPComp'' & ``Lang'' & ``Union'' & ``Optional'' & $p$ \\ 
  \hline
  1 & 1 & 1 & 1 & 1 & 0.012 \\ 
  1 & 1 & 1 & 1 & 2 & $1.4 \cdot 10^{-09}$ \\ 
  1 & 1 & 1 & 2 & 1 & $3.1 \cdot 10^{-09}$ \\ 
  1 & 1 & 1 & 2 & 2 & $6.1 \cdot 10^{-11}$ \\ 
  1 & 1 & 2 & 1 & 1 & $3.3 \cdot 10^{-06}$ \\ 
  1 & 1 & 2 & 1 & 2 & $2.7 \cdot 10^{-09}$ \\ 
  1 & 1 & 2 & 2 & 1 & $2 \cdot 10^{-06}$ \\ 
  1 & 1 & 2 & 2 & 2 & $3.6 \cdot 10^{-10}$ \\ 
  1 & 2 & 1 & 1 & 1 & 0.014 \\ 
  1 & 2 & 1 & 1 & 2 & $1.2 \cdot 10^{-10}$ \\ 
  1 & 2 & 1 & 2 & 1 & $2.8 \cdot 10^{-14}$ \\ 
  1 & 2 & 1 & 2 & 2 & $4.1 \cdot 10^{-15}$ \\ 
  1 & 2 & 2 & 1 & 1 & $2.1 \cdot 10^{-05}$ \\ 
  1 & 2 & 2 & 1 & 2 & $2.7 \cdot 10^{-07}$ \\ 
  1 & 2 & 2 & 2 & 1 & 0.0072 \\ 
  1 & 2 & 2 & 2 & 2 & $1.6 \cdot 10^{-05}$ \\ 
  2 & 1 & 1 & 1 & 1 & 0.28 \\ 
  2 & 1 & 1 & 1 & 2 & $3 \cdot 10^{-07}$ \\ 
  2 & 1 & 1 & 2 & 1 & $3.3 \cdot 10^{-07}$ \\ 
  2 & 1 & 1 & 2 & 2 & $1.7 \cdot 10^{-08}$ \\ 
  2 & 1 & 2 & 1 & 1 & 0.0023 \\ 
  2 & 1 & 2 & 1 & 2 & $6.5 \cdot 10^{-07}$ \\ 
  2 & 1 & 2 & 2 & 1 & 0.00032 \\ 
  2 & 1 & 2 & 2 & 2 & $1.3 \cdot 10^{-06}$ \\ 
  2 & 2 & 1 & 1 & 1 & 0.013 \\ 
  2 & 2 & 1 & 1 & 2 & $1 \cdot 10^{-11}$ \\ 
  2 & 2 & 1 & 2 & 1 & $3.8 \cdot 10^{-11}$ \\ 
  2 & 2 & 1 & 2 & 2 & $4.1 \cdot 10^{-15}$ \\ 
  2 & 2 & 2 & 1 & 1 &   1 \\ 
  2 & 2 & 2 & 1 & 2 & $2.3 \cdot 10^{-05}$ \\ 
  2 & 2 & 2 & 2 & 1 &   1 \\ 
  2 & 2 & 2 & 2 & 2 & 0.99 \\ 
   \hline
\end{tabular}
\end{center}
\end{table}

@@

\begin{figure}[h!]
  \centerline{%
    \includegraphics[width=.9\textwidth]{frac32normal.pdf}}
  \caption{A normal plot of the Fractional Factorial Experiment with
    32 runs. The labelled points are considered significant by using
    the Lenth criterion at a level
    $\alpha=0.05$.}
  \label{fig:frac32normal}
\end{figure}

@@

\begin{figure}[h!]
  \centerline{%
    \includegraphics[width=0.9\textwidth]{frac64normal.pdf}}
  \caption{A normal plot of the Fractional Factorial Experiment with
    64 runs. The labelled points are considered significant by using
    the Lenth criterion. Triangles correspond to a level $\alpha=0.15$
    and crosses at $\alpha=0.05$. The use of different symbols have
    been added to FrF2 by the authors.}\label{fig:frac64normal}
\end{figure}

@@

\begin{figure}[h!]
  \centerline{%
  \includegraphics[width=.9\textwidth]{frac64hugenormal.pdf}}
  \caption{A normal plot of the Fractional Factorial Experiment with
    64 runs where the level 2 of ``TripleC'' is 15 MTriples. The
    labelled points are considered significant by using the Lenth
    criterion at a level $\alpha=0.05$.}\label{fig:frac64hugenormal}
\end{figure}
